A catalog of wide binaries and comoving star pairs from \textsl{Gaia DR1} and low-resolution spectroscopy

### 2017-07-05 paper/comoving-followup.tex
words about new figures

For chemical studies, open clusters have the advantage of having many stars per
cluster, but there are many fewer clusters than comoving stars.
Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
surveys (cite rave, etc.), several hundred such systems are already known
(\citealt{Shaya:2011}).
In addition, over ten-thousand \emph{candidate} comoving stars have been
discovered from astrometric data alone, using data from both the
\project{Hipparcos} mission (\citealt{ESA:1997}---\citealt{Shaya:2011}) and
\tgas\ catalog (\citealt{Gaia-Collaboration:2016a}---
\citealt{Oh:2017,Andrews:2017}).
Radial velocity follow-up is necessary to confirm pairs using these astrometric
catalogs, though \gaia\ \DR{2} will provide radial velocities for the brightest
stars ($G < 12$).
High-resolution spectroscopy is still needed to measure detailed chemical
abundances of these pairs, which may present interesting surprises; one
large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
high-resolution, high signal-to-noise spectra display large abundance
differences ($\approx 0.2~\dex$) that appear to have resulted from massive
accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
submitted).

@@

Of the 323 observed star pairs, 11 have at least one member with an unreliable
(flagged) radial velocity measurement from the pipeline above.
For the remaining 312 pairs, we use the measured radial velocities to re-compute
the marginal likelihood ratio for the two hypotheses considered in previous work
(see Appendix~B, \citealt{Oh:2017}).
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
As with previous work, we set the small dispersion to the orbital velocity of a
binary with the separation of the pair, and use a mixture of three Gaussian
velocity components for the prior, meant to represent the thin disk, thick disk,
and halo velocity distributions relative to the sun.

@@

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information likely supports the belief that the two
stars are truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\, and the log-FML-ratio including the radial velocity data as
\llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 125 of
the 312 comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr}

@@

TODO: check on ipython notebook re-running with updated observation shite

\subsection{Physical properties of genuine comoving pairs}\label{sec:}

\todo{Plot of magnitude of 3d velocity difference vs. 3d separation?}

\todo{separation distribution - compare with Shaya and Olling sample!}
\todo{10 overlap with Shaya and Olling}

\todo{distribution of orbital poles in Galactic coordinates}

@@

\section{Conclusions }

>>>>

For chemical studies, open clusters have the advantage of having many stars per
cluster, but there are many fewer clusters than comoving stars.
Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
surveys (cite rave, etc.), several hundred such systems are already known
(\citealt{Shaya:2011}).
In addition, over ten-thousand \emph{candidate} comoving stars have been
discovered from astrometric data alone, using data from both the
\project{Hipparcos} mission (\citealt{ESA:1997}---\citealt{Shaya:2011}) and
\tgas\ catalog (\citealt{Gaia-Collaboration:2016a}---
\citealt{Oh:2017,Oelkers:2017,Andrews:2017}).
Radial velocity follow-up is necessary to confirm pairs using these astrometric
catalogs, though \gaia\ \DR{2} will provide radial velocities for the brightest
stars ($G < 12$).
High-resolution spectroscopy is still needed to measure detailed chemical
abundances of these pairs, which may present interesting surprises; one
large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
high-resolution, high signal-to-noise spectra display large abundance
differences ($\approx 0.2~\dex$) that appear to have resulted from massive
accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
submitted).

@@

Of the 323 observed star pairs, 11 have at least one member with an unreliable
(flagged) radial velocity measurement from the pipeline above.
For the remaining 311 pairs, we use the measured radial velocities to re-compute
the marginal likelihood ratio for the two hypotheses considered in previous work
(see Appendix~B, \citealt{Oh:2017}).
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
As with previous work, we set the small dispersion to the orbital velocity of a
binary with the separation of the pair, and use a mixture of three Gaussian
velocity components for the prior, meant to represent the thin disk, thick disk,
and halo velocity distributions relative to the sun.

@@

@@

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information likely supports the belief that the two
stars are truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\, and the log-FML-ratio including the radial velocity data as
\llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 128 of
the 311 comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio for the two stars in all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution over velocity-difference as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.

\subsection{Physical properties of genuine comoving pairs}\label{sec:separation}

\figurename~\ref{fig:dx-dv} shows the magnitude of the 3D velocity difference
plotted against the 3D positional separation of the two stars in all observed
pairs.
We again plot the median of the velocity difference distribution, and here use
the Lutz-Kelker-corrected distance estimate (\citealt{Lutz:1973}) computed from
the stars' \tgas\ parallax measurements to compute the 3D positions of the two
stars.
Small, grey markers are pairs for which $\llrnew < \llrold$, i.e. where the
radial velocity measurements weaken or change our belief that a given pair is
comoving.
Larger, black markers are genuine comoving pairs ($\llrnew > \llrold$); note
that many large-separation pairs have been ruled-out as comoving, but
most small-separation pairs maintain their status as comoving.
This is also apparent in \figurename~\ref{fig:separation}, which shows
histograms of the spatial separation and sky separation for all pairs;
background, grey histogram shows all observed pairs, and foreground, black
histogram shows all pairs for which $\llrnew > \llrold$.

At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
genuinely comoving; interestingly, the separation distribution of the genuine
pairs appears to be at least uniform or possibly growing in number as the
separation increases.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{dx-dv.pdf}
  \end{center}
  \caption{%
    FML
    \label{fig:dx-dv}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    FML
    \label{fig:separation}}
\end{figure}

\todo{separation distribution - compare with Shaya and Olling sample!}
\todo{10 overlap with Shaya and Olling}
\todo{RAVE pairs - both observed}

@@

\section{Conclusions}
